Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-11-04 manuscript.tex
general intro comments adjusted

% Migrant Example Pathways
% Relevance (Version 1 of 2): Migrant Example Version:
Conflict between social groups and their individual members remains a prevalent feature of the modern human condition. One of the main examples to date are the struggles of many migrants across the world, hoping to build a positive relationship with the majority group. One of the main ameliorations proposed has been intergroup contact. In its most essential interpretation, the intergroup contact hypothesis postulates that prejudice can be reduced and favorable attitudes increased if members of two groups have frequent and positive contact \citep[e.g.,][]{Allport1954b, Hewstone1996, Pettigrew1998}. Over the past 70 years, a plethora of studies and interventions have shown the general effectiveness of positive intergroup contact \citep[e.g.,][]{Pettigrew2006}. However, even though a central assumption of intergroup contact theory is that the contact should be positive, relatively little research has thus far explained when and why people perceive their everyday inter-group interactions as positive.

@@

% Problem v.02: theoretical interaction quality central to understanding outcomes, practical many impactful negative interactions in everyday life
% Problem Illustration Paragraph:
Importantly, not understanding when and why an interaction is perceived as positive presents substantial theoretical and practical obstacles. There is now consistent evidence that negative intergroup contacts lead to worse attitudes, prejudice, and reduced future interaction motivation \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}. Theoretically, understanding interaction quality (i.e., whether an interaction is perceived as positive or negative) thus sits at the heart of understanding when an intergroup contact is successful (e.g., Brown et al. \citeyear{Brown2007}, Tropp et al. \citeyear{Tropp2016}; also cf. Allport’s original sentiment, Allport \citeyear{Allport1954b}). Practically, policymakers and practitioners are thus far often under-prepared to deal with the occurrences of negative interactions, especially in every-day life contexts. Understanding the psychological mechanisms of when and why interactions are perceived as positive is, thus, an important issue for understanding whether an interaction leads to better intergroup perceptions, especially during everyday interactions outside the lab.

% Aim / Solution v.02: look at need fulfillment as mechanism in daily interactions of migrants
% Proposal Paragraph:
In this article, we turn towards the motivational literature and propose that the fulfillment of fundamental psychological needs might offer a psycho-social mechanism for understanding and explaining contact quality and positive outgroup attitudes. That is to say, that if a person, for example, seeks to feel accepted by their interaction partner and this need is fulfilled during the interaction, the person should rate the interaction and the group of the interaction partner more favorably. To test this, we collected three sets of real-life data from recent immigrants. More specifically, we followed their daily interactions with majority group members, tracking situational needs, interaction quality, and outgroup attitudes.

\section{Need Mechanism in Intergroup Contact}
% Motivational mechanism:
% interaction quality is important
% past research has either focused on conditions or cognitive-affective processes
% neither are necessary or explain why an interaction is positive
Several meta-analytic reviews have found that positive intergroup contact reduces prejudice and increases positive attitudes in experimental and cross-sectional studies \citep[][]{Tropp2005, Pettigrew2006, Davies2011}, as well as in intergroup contact interventions outside the lab \citep[][]{Beelmann2014, Lemmer2015}. It is widely accepted that equal status, common goals, collaboration, and structural support during the interaction form the optimal conditions for such positive contact effects \citep[Allport's Optimal Contact conditions;][]{Allport1954b, Pettigrew1969}. However, a major meta-analysis of these contact conditions showed that contact resulted in more positive intergroup relations even when Allport's conditions were not met \citep[][]{Pettigrew2006}. It, thus, remains unclear when and why exactly an interaction is perceived as positive and what factors (beyond Allport's conditions) explain positive contact effects\footnote{It should be noted here that following Allport's original conditions, several additional conditions of optimal contact were proposed \citep[for a critical discussion see][]{Pettigrew1986}. Similarly, since the early 21\textsuperscript{st} century many psychological processes during intergroup contact were examined \citep[e.g. see,][]{Paolini2021}. Among others, researchers have explored different forms of social categorizations \citep[][]{Pettigrew1998}, the salience of social categories \citep[][]{Brown2005}, intimacy \citep[e.g.,][]{Marinucci2021} and attachment \citep[e.g.,][]{Tropp2021}, threat and intergroup anxiety \citep[e.g.,][]{Stephan2008, Paolini2004}, and to a lesser extent knowledge about the other group \citep[][]{Pettigrew2008c}. However, all these advances share the underlying criticism that they are not geared towards understanding when and why an interaction is perceived as positive.}. We propose that we may look to the literature on motivation and psychological needs to understand when and why exactly an interaction is perceived as positive.

% ALTERNATIVE: SHORTER BUT NOT FOCUSED ON ALLPORT'S CONDITIONS (WHICH WE TEST LATER)
% Several meta-analytic reviews found that positive intergroup contact reduces prejudice and increases positive attitudes in experimental and cross-sectional studies \citep[][]{Tropp2005, Pettigrew2006, Davies2011}, as well as in intergroup contact interventions outside the lab \citep[][]{Beelmann2014, Lemmer2015}. Past research has either focused on conditions under which interactions tend to have favorable effects \citep[e.g., Allport's optimal conditions;][]{Allport1954b, Pettigrew1969} or has investigated cognitive-affective processes during contact \citep[e.g., group salience, intimacy, attachment, or intergroup anxiety;][]{Brown2005, Marinucci2021, Tropp2021, Paolini2004}. However, neither the contact conditions nor the cognitive-affective processes are particularly well equipped to explain when and why an interaction will be perceived as positive. We propose that we may look to the literature on motivation and psychological needs to understand when and why exactly an interaction is perceived as positive.

% Important in broader field: needs have been found to be important in broader intergroup relations and interpersonal contact literature.
Although considerations of psychological needs and motivations have remained markedly absent within the intergroup contact literature, the concepts are in no way foreign to the broader topic of intergroup relations. Psychological needs have for example found fruitful utility in reconciliation research \citep[][]{Shnabel2008}, the study of minority's discrimination experiences \citep[][]{Celebi2017}, and in non-intergroup interactions \citep[][]{Downie2008}. 
% On the most general level, research on intergroup conflicts has, for example, shown that addressing the group-specific needs predicts willingness to reconcile with the other group \citep[][]{Shnabel2008}. For the case of refugee migrants in their relationship with a majority group, a recent study found that identity needs will buffer against discrimination experiences and will protect personal health \citep[][]{Celebi2017}. Additionally, within the field of interpersonal relations, we find first evidence that the fulfillment of fundamental psychological needs might predict perceived interaction quality in non-intergroup interactions \citep[][]{Downie2008}. 
% Promising for intergroup contact: Asked for in intergroup contact reviews and first research looked at social change
And even within the field of intergroup contact the idea of need fulfillment as a psychological mechanism is not entirely novel. \citet{Dovidio2017} in their narrative review have suggested that understanding psychological needs might be essential to constructive intergroup contact. And recently first empirical research has even found psychological need fulfillment during intergroup contacts to predict support for social change \citep[][]{Hassler2021}. There is thus initial evidence to suggest that need fulfillment during intergroup contact might offer a promising psychological mechanism to understand perceived interaction quality and positive outgroup attitudes.

% Difficult to test general mechanism: Either consider too many needs (situationally relevant but not feasible) or too few (feasible but not transferable). Solution: ask for main need + rating of main need  
% MAYBE MOVE THIS TO 'The Present Research Section'
One reason why motivational considerations might have remained absent from the intergroup contact literature is that there is an overwhelming number of individual needs or goals that might be relevant to a person during an intergroup interaction. Researchers considering the motivational content would, thus, either test few hyper-specific needs that might not be transferable to other intergroup contexts or they may need to assess a broad range of fundamental psychological needs. To avoid this predicament we propose to start by using adaptive and responsive survey designs that allow a tailored approach based on the participants inputs \citep[e.g.,][]{Tourangeau2017}. In particular, we propose to ask the participants to report their main goal during the interaction in a short open-ended question (i.e., situation core need) and with reference to their own response the participants can then indicate how much this need was fulfilled during the interaction (i.e., need fulfillment). Such an adaptive approach would allow us to take this initial step and test whether situational psychological needs indeed generally predict perceived interaction quality and positive outgroup attitudes independent of need content.

\section{Real-Life Process Data}
While we have argued that a need fulfillment mechanism is relevant to intergroup contact generally, its flexible and broad applicability might be ideally suited to address the pressing issue of understanding natural intergroup contacts outside the lab. Investigations of such `real-life interactions' often suffer from the difficulty that past intergroup contact research has either focused on the mechanisms of individual interactions in artificial lab studies or has focused on longer-term recall self-reports of natural interactions \citep[e.g.,][]{Pettigrew2006}. Even with extended intervention studies, the most fine-grained data available is usually limited to pre-post-control designs. Such empirical practices stand, however, in stark contrast to many of the theoretical advances that have focused on the dynamic nature of intergroup relations \citep[e.g.,][]{Pettigrew1998}, as well as the core idea of the contact hypothesis, which was focused on the daily interactions of people \citep[see][]{Allport1954b}. As a result, prominent figures within the intergroup contact field have called for studies that collect longitudinal \citep[][]{Pettigrew1998, Pettigrew2008, Pettigrew2008b, Pettigrew2011} and real-life experience-sampling data outside the lab \citep[][]{MacInnis2015, McKeown2017, Dixon2005}. Such data would be able to capture real-life interactions that include interaction-specific mechanism information close the actual experience\footnote{Additionally, such experience-sampling data can be collected close to the intergroup interactions and would, thus, largely mitigate recall biases. Moreover, because data is nested within participants, experience-sampling data often allows to capture large amounts of high-quality data with relatively few participants.}.

% Used to be difficult but technological and methodological developments (e.g., FormR + HLM) → collect large body of intensive longitudinal data
In the past, such data collections were often unfeasible because they were either physically impractical or too expensive. However, recent technological developments allow us to easily collect experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}. At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable \citep[e.g., see][]{ODonnell2021}. Given these, technological and methodological developments, we were able to collect a large amount of real-life data following the daily intergroup interactions of recent migrants with the majority group members.

\section{The Present Research}
The aim of this paper is essentially threefold. First, our study is among the first to test the fundamental tenets of intergroup contact and Allport's conditions in real-life extensive longitudinal data. Second, we test whether the fulfillment situational psychological needs is meaningfully related to more positive outgroup attitudes following intergroup interactions. And third, we test the role of interaction quality in  the relationship of psychological need fulfillment and outgroup attitudes. Based on this we formulated three main hypotheses:
\begin{enumerate}[label=H\arabic*:]
    \item Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes across extensive longitudinal data.
    \item Based on Allport’s optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data.
    \item Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data. We also expect situational need fulfillments to work at least as good as Allport's optimal contact conditions in predicting outgroup attitudes.
\end{enumerate}
Beyond these main hypotheses we have formulated several sub-hypotheses that form the basis for our analysis plan (these sub-hypotheses also include several robustness checks and are available in Appendix \ref{app:AppendixHypotheses}).

@@

\hl{[link sub-hypotheses to the following?]}\\
\hl{[add formulas to the following?]}\\
\hl{[move (parts of) the following to Appendix }\ref{app:AppendixHypotheses}\hl{?]}\\
\hl{[I think this is too long]}

@@

\begin{figure}
  \caption{Robustness Analyses}
  \label{fig:Robustness}
  \centering\includegraphics[width=\textwidth]{Figures/forestParametricRobustnessComb.png}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.65\textwidth]{Figures/forestParametricFERobustContact.png}
  %\end{subfigure}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.65\textwidth]{Figures/forestParametricFERobustSDT.png}
  %\end{subfigure}
  \caption*{Note: \\
  (a) Need Fulfillment and Intergroup Contact predicting Outgroup Attitudes (full sample).\\
  (b) Core Need Fulfillment predicting Outgroup Attitudes, while controlling for self-determination theory needs (intergroup contact sample).\\
  General: Random effects meta analytic results are presented for completeness.}
\end{figure}

@@

\begin{subhyp}[H\ref{hyp:contactFreq}] \label{hyp:contactFreq}
\addtolength{\leftskip}{2.5em}
Participants with more intergroup interactions should have a more favorable outgroup attitudes.
\end{subhyp}

@@

\begin{subhyp}[H\ref{hyp:contactFreqQual}] \label{hyp:contactFreqQual}
\addtolength{\leftskip}{2.5em}
Participants with more intergroup interactions should have a more favorable outgroup attitudes depending on the average interaction quality.
\end{subhyp}

@@

\begin{subhyp}[H\ref{hyp:AllportsQualityMediation}] \label{hyp:AllportsQualityMediation}
\addtolength{\leftskip}{2.5em}
Based on past research on the role of interaction quality, the variance explained in outgroup attitudes by Allport's optimal contact should to a large extend be assumed by interaction quality.
\end{subhyp}

@@

\setcounter{subhyp}{0}
\begin{subhyp}[H\ref{hyp:keyNeedPred}] \label{hyp:keyNeedPred}
\addtolength{\leftskip}{2.5em}
Outgroup attitudes should be more favorable after intergroup interactions with high key need fulfillment.
\end{subhyp}

@@

\begin{subhyp}[H\ref{hyp:keyNeedContactType}] \label{hyp:keyNeedContactType}
\addtolength{\leftskip}{2.5em}
The effect of key need fulfillment on outgroup attitudes should be specific to intergroup interactions and not be due to need fulfillment in general. Thus, the effect of key need fulfillment on outgroup attitudes should stronger for intergroup interact than for ingroup interactions. 
\end{subhyp}

\begin{subhyp}[H\ref{hyp:keyNeedSDT}] \label{hyp:keyNeedSDT}
\addtolength{\leftskip}{2.5em}
The effect of key need fulfillment on outgroup attitudes should be persist even when taking other fundamental psychological needs into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for autonomy, competence, and relatedness fulfillment during the interaction (cf., self-determination theory). 
\end{subhyp}

@@

>>>>

% Migrant Example Pathways
% Relevance (Version 1 of 2): Migrant Example Version:
One of the main intergroup societal issues to date, are the struggles of many migrants across the world, hoping to build a positive relationship with the majority group. The intergroup contact hypothesis postulates that prejudice can be reduced and favorable attitudes be increased if members of two groups have frequent and positive contact \citep[e.g.,][]{Allport1954b, Hewstone1996, Pettigrew1998}. Over the past 70 years, a plethora of studies and interventions have shown the general effectiveness of positive intergroup contact \citep[e.g.,][]{Pettigrew2006}. However, even though a central assumption of intergroup contact theory is that the contact should be positive, relatively little research has thus far explained when and why people perceive their everyday intergroup interactions as positive.

@@

% Problem v.02: theoretical interaction quality central to understanding outcomes, practical many impactful negative interactions in everyday life
% Problem Illustration Paragraph:       
Importantly, as we still fail to understand when and why an interaction is perceived as positive, substantial theoretical and practical challenges remain. There is now consistent evidence that negative intergroup contacts lead to worse attitudes, prejudice, and reduced future interaction motivation \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}. In light of these findings, understanding interaction quality thus sits at the heart of understanding when an intergroup contact is successful (e.g., Brown et al. \citeyear{Brown2007}, Tropp et al. \citeyear{Tropp2016}; also cf. Allport’s original sentiment, Allport \citeyear{Allport1954b}). Practically, policymakers and practitioners are thus far often under-prepared to deal with the occurrences of negative interactions, especially in everyday life contexts. Understanding the psychological mechanisms of when and why interactions are perceived as positive is, thus, an important issue for understanding whether an interaction leads to better intergroup perceptions, especially during everyday interactions outside the lab.

% Aim / Solution v.02: look at need fulfillment as a mechanism in daily interactions of migrants
% Proposal Paragraph:
We propose that the key to understanding how an interaction is perceived is to examine the level of need fulfillment it provided to an individual. As an example, if someone seeks acceptance by their interaction partner, and this need is fulfilled during the interaction, the person should rate the interaction and the group of the interaction partner more favorably. To test this, we collected three sets of real-life data from recent immigrants, assessing their daily interactions with majority group members, tracking situational needs, interaction quality, and outgroup attitudes.

\section{Need Mechanism in Intergroup Contact}
% Motivational mechanism:
% interaction quality is important
% past research has either focused on conditions or cognitive-affective processes
% neither are necessary nor explain why an interaction is positive
Several meta-analytic reviews have found that positive intergroup contact reduces prejudice and increases positive attitudes in experimental and cross-sectional studies \citep[][]{Tropp2005, Pettigrew2006, Davies2011}, as well as in intergroup contact interventions outside the lab \citep[][]{Beelmann2014, Lemmer2015}. It is widely accepted that equal status, common goals, collaboration, and structural support during the interaction form the optimal conditions for such positive contact effects \citep[][]{Allport1954b, Pettigrew1969}. However, despite their prominence in guiding research on this topic, meeting the contact conditions does not seem to be essential to finding positive effects of intergroup contact \citep[][]{Pettigrew2006}. It, thus, remains unclear when and why an interaction is perceived as positive and what factors (beyond Allport's conditions) explain positive contact effects\footnote{It should be noted here that following Allport's original conditions, several additional conditions of optimal contact were proposed \citep[for a critical discussion see][]{Pettigrew1986}. Similarly, since the early 21\textsuperscript{st} century many psychological processes during intergroup contact were examined \citep[e.g. see,][]{Paolini2021}. Among others, researchers have explored different forms of social categorizations \citep[][]{Pettigrew1998}, the salience of social categories \citep[][]{Brown2005}, intimacy \citep[e.g.,][]{Marinucci2021} and attachment \citep[e.g.,][]{Tropp2021}, threat and intergroup anxiety \citep[e.g.,][]{Stephan2008, Paolini2004}, and to a lesser extent knowledge about the other group \citep[][]{Pettigrew2008c}. However, all these advances share the underlying criticism that they are not geared towards understanding when and why an interaction is perceived as positive.}. We propose that the literature on motivation and psychological needs may help us understand when and why exactly an interaction is perceived as positive.

% ALTERNATIVE: SHORTER BUT NOT FOCUSED ON ALLPORT'S CONDITIONS (WHICH WE TEST LATER)
% Several meta-analytic reviews found that positive intergroup contact reduces prejudice and increases positive attitudes in experimental and cross-sectional studies \citep[][]{Tropp2005, Pettigrew2006, Davies2011}, as well as in intergroup contact interventions outside the lab \citep[][]{Beelmann2014, Lemmer2015}. Past research has either focused on conditions under which interactions tend to have favorable effects \citep[e.g., Allport's optimal conditions;][]{Allport1954b, Pettigrew1969} or has investigated cognitive-affective processes during contact \citep[e.g., group salience, intimacy, attachment, or intergroup anxiety;][]{Brown2005, Marinucci2021, Tropp2021, Paolini2004}. However, neither the contact conditions nor the cognitive-affective processes are particularly well-equipped to explain when and why an interaction will be perceived as positive. We propose that we may look to the literature on motivation and psychological needs to understand when and why exactly an interaction is perceived as positive.

% Important in broader field: needs have been found to be important in broader intergroup relations and interpersonal contact literature.
Although considerations of psychological needs and motivations have remained markedly absent within the intergroup contact literature, the concepts are in no way foreign to the broader topic of intergroup relations. Psychological needs have for example found fruitful utility in reconciliation research \citep[][]{Shnabel2008}, the study of minority's discrimination experiences \citep[][]{Celebi2017}, and in non-intergroup interactions \citep[][]{Downie2008}. 
% On the most general level, research on intergroup conflicts has, for example, shown that addressing the group-specific needs predicts willingness to reconcile with the other group \citep[][]{Shnabel2008}. For the case of refugee migrants in their relationship with a majority group, a recent study found that identity needs will buffer against discrimination experiences and will protect personal health \citep[][]{Celebi2017}. Additionally, within the field of interpersonal relations, we find first evidence that the fulfillment of fundamental psychological needs might predict perceived interaction quality in non-intergroup interactions \citep[][]{Downie2008}. 
% Promising for intergroup contact: Asked for in intergroup contact reviews and first research looked at social change
Even within the field of intergroup contact, the idea of need fulfillment as a psychological mechanism is not entirely novel \citep[e.g.,][]{Dovidio2017}. First empirical research has even found psychological need fulfillment during intergroup contacts to predict support for social change \citep[][]{Hassler2021}. Considering need fulfillment during intergroup contact might therefore offer a promising psychological mechanism to understand perceived interaction quality and positive outgroup attitudes.

% Difficult to test general mechanism: Either consider too many needs (situationally relevant but not feasible) or too few (feasible but not transferable). Solution: ask for main need + rating of main need  
% MAYBE MOVE THIS TO 'The Present Research Section'
One reason why motivational considerations might have remained absent from the intergroup contact literature is that there is an overwhelming number of individual needs or goals that might be relevant to a person during an intergroup interaction. Researchers considering the motivational content would, thus, either test few hyper-specific needs that might not be transferable to other intergroup contexts or they may need to assess a broad range of fundamental psychological needs. To avoid this predicament we propose to start by using adaptive and responsive survey designs that allow a tailored approach based on the participants' inputs \citep[e.g.,][]{Tourangeau2017}. In particular, we propose to ask the participants to report their main goal during the interaction in a short open-ended question (i.e., situation core need) and with reference to their own response the participants can then indicate how much this need was fulfilled during the interaction (i.e., need fulfillment). Such an adaptive approach allows us to take this initial step and test whether situational psychological needs indeed generally predict perceived interaction quality and positive outgroup attitudes independent of need content.

\section{Intergroup Contact in Daily Life}
While we have argued that a need fulfillment mechanism is relevant to intergroup contact generally, its flexible and broad applicability might be ideally suited to address the pressing issue of understanding natural intergroup contacts outside the lab. Investigations of such `real-life interactions' often suffer from the difficulty that past intergroup contact research has either focused on the mechanisms of individual interactions in artificial lab studies or has focused on longer-term recall self-reports of natural interactions \citep[e.g.,][]{Pettigrew2006}. Even with extended intervention studies, the most fine-grained data available is usually limited to pre-post-control designs. Such empirical practices stand, however, in stark contrast to many of the theoretical advances that have focused on the dynamic nature of intergroup relations \citep[e.g.,][]{Pettigrew1998}, as well as the core idea of the contact hypothesis, which was focused on the daily interactions of people \citep[][]{Allport1954b}. As a result, within the intergroup contact field researchers have called for longitudinal \citep[][]{Pettigrew1998, Pettigrew2008, Pettigrew2008b, Pettigrew2011} and real-life experience-sampling data outside the lab \citep[][]{MacInnis2015, McKeown2017, Dixon2005}. Such data would be able to capture real-life interactions that include interaction-specific mechanism information close to the actual experience\footnote{Additionally, such experience-sampling data can be collected close to the intergroup interactions and would, thus, largely mitigate recall biases. Moreover, because data is nested within participants, experience-sampling data often allows capturing large amounts of high-quality data with relatively few participants.}.

% Used to be difficult but technological and methodological developments (e.g., FormR + HLM) → collect large body of intensive longitudinal data
In the past, such data collections were often unfeasible because they were either physically impractical or too expensive. However, recent technological developments allow us to easily collect experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}. At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable \citep[e.g., see][]{ODonnell2021}. Given these technological and methodological developments, we were able to collect a large amount of real-life data following the daily intergroup interactions of recent migrants with the majority group members.

\section{The Present Research}
The aim of this paper is essentially threefold. First, our study is among the first to test the fundamental tenets of intergroup contact and Allport's conditions in real-life extensive longitudinal data. Second, we test whether the fulfillment of situational psychological needs is meaningfully related to more positive outgroup attitudes following intergroup interactions. And third, we test the role of interaction quality in  the relationship of psychological need fulfillment and outgroup attitudes. Based on this we formulated three main hypotheses:
\begin{enumerate}[label=H\arabic*:]
    \item Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes across extensive longitudinal data.
    \item Based on Allport’s optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data.
    \item Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data. We also expect situational need fulfillments to work at least as well as Allport's optimal contact conditions in predicting outgroup attitudes.
\end{enumerate}
Beyond these main hypotheses, we have formulated several sub-hypotheses that form the basis for our analysis plan (these sub-hypotheses also include several robustness checks and are available in Appendix \ref{app:AppendixHypotheses}).

@@

@@

\begin{figure}
  \caption{Robustness Analyses}
  \label{fig:Robustness}
  \centering\includegraphics[width=\textwidth]{Figures/forestParametricRobustnessComb.png}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.65\textwidth]{Figures/forestParametricFERobustContact.png}
  %\end{subfigure}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.65\textwidth]{Figures/forestParametricFERobustSDT.png}
  %\end{subfigure}
  \caption*{Note: \\
  (a) Need Fulfillment and Intergroup Contact predicting Outgroup Attitudes (full sample).\\
  (b) Core Need Fulfillment predicting Outgroup Attitudes while controlling for self-determination theory needs (intergroup contact sample).\\
  General: Random effects meta-analytic results are presented for completeness.}
\end{figure}

@@

\begin{subhyp}[H\ref{hyp:contactFreq}] \label{hyp:contactFreq}
\addtolength{\leftskip}{2.5em}
Participants with more intergroup interactions should have more favorable outgroup attitudes.
\end{subhyp}

@@

\begin{subhyp}[H\ref{hyp:contactFreqQual}] \label{hyp:contactFreqQual}
\addtolength{\leftskip}{2.5em}
Participants with more intergroup interactions should have more favorable outgroup attitudes depending on the average interaction quality.
\end{subhyp}

@@

\begin{subhyp}[H\ref{hyp:AllportsQualityMediation}] \label{hyp:AllportsQualityMediation}
\addtolength{\leftskip}{2.5em}
Based on past research on the role of interaction quality, the variance explained in outgroup attitudes by Allport's optimal contact should to a large extent be assumed by interaction quality.
\end{subhyp}

@@

\setcounter{subhyp}{0}
\begin{subhyp}[H\ref{hyp:keyNeedPred}] \label{hyp:keyNeedPred}
\addtolength{\leftskip}{2.5em}
Outgroup attitudes should be more favorable after intergroup interactions with high key-need fulfillment.
\end{subhyp}

@@

\begin{subhyp}[H\ref{hyp:keyNeedContactType}] \label{hyp:keyNeedContactType}
\addtolength{\leftskip}{2.5em}
The effect of key need fulfillment on outgroup attitudes should be specific to intergroup interactions and not be due to need fulfillment in general. Thus, the effect of key need fulfillment on outgroup attitudes should be stronger for intergroup interactions than for ingroup interactions. 
\end{subhyp}

\begin{subhyp}[H\ref{hyp:keyNeedSDT}] \label{hyp:keyNeedSDT}
\addtolength{\leftskip}{2.5em}
The effect of key need fulfillment on outgroup attitudes should persist even when taking other fundamental psychological needs into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for autonomy, competence, and relatedness fulfillment during the interaction (cf., self-determination theory). 
\end{subhyp}

@@

\section{Appendix B: Robustness}
\label{app:AppendixRobustness}
